---
title: "Statistical Inference Project Part 2"
Date: Dec. 21st. 2014
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    theme: spacelab
    toc: yes
  pdf_document:
    fig_caption: yes
    highlight: tango
    number_sections: yes
    toc: yes
Author: Ahmed Assal
---
## Synopsis
This is the project for the statistical inference class.  This is the second part of the project which is an analysis of the ToothGrowth data. 

In the second portion of the class, I am going to analyze the ToothGrowth data in the R datasets package.

## Data Description    
The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

[,1]   len   numeric	 Tooth length   
[,2]	 supp	 factor	 Supplement type (VC or OJ).   
[,3]	 dose	 numeric	 Dose in milligrams.   
*(source: R Doumentation)*  

## Data Processing
### Loading packages and libraries   
```{r results='hide'}
library(ggplot2)
library(datasets)
library(reshape2)
library(dplyr)
```
### Loading dataset
```{r}
toothGrowth = tbl_df(ToothGrowth)
```

## Anaylsis
### Summarizing Data

```{r}
glimpse(toothGrowth)
summary(toothGrowth)
head(toothGrowth,3); tail(toothGrowth,3)
```

### Exploratory Data Analysis
#### Tooth length vs supplement type

Fisrt, I will explore the relaion between the supplement type and the lenth of tooth
    
```{r}
ggplot(data=toothGrowth, aes(x=len)) + 
  geom_histogram(binwidth=2) +
  facet_grid( . ~ supp) +
  xlab("Tooth Length") +
  ylab("Count")
```   

It seems that both supplements have similar effects on the length of tooth, however, the orange juice, OJ, has more profound effect on most of the teeth    

#### Tooth length vs dosage    
I will then explore the effect of the dosage quantity on the length of the tooth while examinning this effect for both of the supplements.    

```{r}
ggplot(data=toothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) + 
  geom_bar(stat="identity") +
  xlab("Dosage (mg)") +
  ylab("Tooth Length") +
  guides(fill=guide_legend(title="Supplement Type"))
```     

As the dosage increases so does the length of the tooth, however, again the orange juice, the OJ supplement has a stronger effect on the tooth length.

### Comparison of tooth growth by supp and dose   
#### Groups based on supplement
##### Extracting Groups
```{r}
oj= subset(toothGrowth, supp=="OJ")$len
vc= subset(toothGrowth, supp=="VC")$len

```    
##### Are groups of different variance ?
```{r}
var(oj);var(vc)

```   
The variance is different for all groups created based on supplement    
##### Comparing the supplement groups using T Test   
```{r}
test= t.test(len ~ supp, paired = FALSE, val.equal = FALSE, data = toothGrowth)
test$conf

```    

#### Groups based on supplement and dosage
##### Extracting Groups
```{r}
oj05= subset(toothGrowth, dose ==0.5 & supp=="OJ")$len
oj1= subset(toothGrowth, dose ==1 & supp=="OJ")$len
oj2= subset(toothGrowth, dose ==2 & supp=="OJ")$len


vc05= subset(toothGrowth, dose ==0.5 & supp=="VC")$len
vc1= subset(toothGrowth, dose ==1 & supp=="VC")$len
vc2= subset(toothGrowth, dose ==2 & supp=="VC")$len

```    
##### Are groups of different variance ?
```{r}
var(oj05);var(oj1);var(oj2)
var(vc05);var(vc1);var(vc2)
```   
The variance is different for all groups created based on supplement and dosage    

#### Groups based on dosage
##### Extracting Groups
```{r}
dose05= subset(toothGrowth, dose ==0.5 )$len
dose1= subset(toothGrowth, dose ==1)$len
dose2= subset(toothGrowth, dose ==2)$len
```    
##### Are groups of different variance ?
```{r}
var(dose05);var(dose1);var(dose2)

```   
The variance is different for all groups created based on dosage    


## Conslusions and Assumptions        
    Did the student describe the assumptions needed for their conclusions?
    - The variance differes among the different groups, as calculated.

    
