---
title: "Statistical Inference Project Part 2"
Date: Dec. 21st. 2014
output:
  pdf_document:
    fig_caption: yes
    highlight: tango
    number_sections: yes
    toc: yes
  html_document:
    fig_caption: yes
    number_sections: yes
    theme: spacelab
    toc: yes
Author: Ahmed Assal
---
## Synopsis
This is the project for the statistical inference class.  This is the second part of the project which is an analysis of the ToothGrowth data. 

In the second portion of the class, I am going to analyze the ToothGrowth data in the R datasets package.

## Data Description    
The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

[,1]   len   numeric	 Tooth length   
[,2]	 supp	 factor	 Supplement type (VC or OJ).   
[,3]	 dose	 numeric	 Dose in milligrams.   
*(source: R Doumentation)*  

## Data Processing
### Loading packages and libraries   
```{r results='hide'}
library(ggplot2)
library(datasets)
library(reshape2)
library(dplyr)
```
### Loading dataset
```{r}
toothGrowth = tbl_df(ToothGrowth)
```

## Anaylsis
### Summarizing Data

```{r}
glimpse(toothGrowth)
summary(toothGrowth)
head(toothGrowth,3); tail(toothGrowth,3)
```

### Exploratory Data Analysis
#### Tooth length vs supplement type

Fisrt, I will explore the relaion between the supplement type and the lenth of tooth
    
```{r}
ggplot(data=toothGrowth, aes(x=len)) + 
  geom_histogram(binwidth=2) +
  facet_grid( . ~ supp) +
  xlab("Tooth Length") +
  ylab("Count")
```   

It seems that both supplements have similar effects on the length of tooth, however, the orange juice, OJ, has more profound effect on most of the teeth    

#### Tooth length vs dosage    
I will then explore the effect of the dosage quantity on the length of the tooth while examinning this effect for both of the supplements.    

```{r}
ggplot(data=toothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) + 
  geom_bar(stat="identity") +
  xlab("Dosage (mg)") +
  ylab("Tooth Length") +
  guides(fill=guide_legend(title="Supplement Type"))
```     

As the dosage increases so does the length of the tooth, however, again the orange juice, the OJ supplement has a stronger effect on the tooth length.

### Comparison of tooth growth by supp and dose   
#### Groups based on supplement
##### Extracting Groups
```{r}
oj= subset(toothGrowth, supp=="OJ")$len
vc= subset(toothGrowth, supp=="VC")$len

```    
##### Are groups of different variance ?
```{r}
var(oj);var(vc)

```   

The variance is different for all groups created based on supplement    

##### Comparing the supplement groups using T Test      
```{r}
test= t.test(len ~ supp, paired = FALSE, val.equal = FALSE, data = toothGrowth)

```    
The p-value is `r test$p.value` and the 95% confidence interval is `r test$conf` which suggests that there is no substantial difference between the tooth growth of those subjected to the two types of supplements, namely, OJ and VC, regardless of the dosage.   

#### Groups based on supplement and dosage
##### Extracting Groups
```{r}
oj05= subset(toothGrowth, dose ==0.5 & supp=="OJ")$len
oj1= subset(toothGrowth, dose ==1 & supp=="OJ")$len
oj2= subset(toothGrowth, dose ==2 & supp=="OJ")$len


vc05= subset(toothGrowth, dose ==0.5 & supp=="VC")$len
vc1= subset(toothGrowth, dose ==1 & supp=="VC")$len
vc2= subset(toothGrowth, dose ==2 & supp=="VC")$len

```    
##### Are groups of different variance ?
```{r}
var(oj05);var(oj1);var(oj2)
var(vc05);var(vc1);var(vc2)
```   
The variance is different for all groups created based on supplement and dosage    

##### Comparing the supplement groups using T Test      
```{r}
testOJ05_1= t.test(x= oj1, y=oj05,  val.equal = FALSE)
testOJ05_1$p.value; testOJ05_1$conf.int

testOJ1_2= t.test(x= oj2, y=oj1,  val.equal = FALSE)
testOJ1_2$p.value; testOJ1_2$conf.int

testOJ05_2= t.test(x= oj2, y=oj05,  val.equal = FALSE)
testOJ05_2$p.value; testOJ05_2$conf.int

```    
The p-values and the 95% confidence intervals suggest that as we increase the dosage the teeth growth increases for both supplements, however, the increase in the tooth growth is more aparent in the increase of the dosage from 0.5 to 1 or from 0.5 to 2. The increase in the dosage from 1 to 2 results in more increase in the tooth growth with the VC supplement more than that of the OJ supplement.   

#### Groups based on dosage
##### Extracting Groups
```{r}
dose05_1= subset(toothGrowth, dose ==0.5|dose ==1)
dose05_1$supp = NULL
dose1_2= subset(toothGrowth, dose ==1|dose ==2 )
dose1_2$supp = NULL
dose05_2= subset(toothGrowth, dose ==0.5|dose ==2)
dose05_2$supp = NULL
```    
##### Are groups of different variance ?
```{r}
var(dose05_1);var(dose1_2);var(dose05_2)

```   
The variance is different for all groups created based on dosage    

##### Comparing the supplement groups using T Test      
```{r}
testD05_1= t.test(len ~ dose, paired = FALSE, val.equal = FALSE, data = dose05_1)
testD05_1$p.value; testD05_1$conf.int

testD1_2= t.test(len ~ dose, paired = FALSE, val.equal = FALSE, data = dose1_2)
testD1_2$p.value; testD1_2$conf.int

testD05_2= t.test(len ~ dose, paired = FALSE, val.equal = FALSE, data = dose05_2)
testD05_2$p.value; testD05_2$conf.int

```    
The p-values and the 95% confidence intervals suggest that there is no substantial difference between the tooth growth of the different dosages regardless of the supplement type.   

## Assumptions and conclusions        
    
    - The variance differes among the different groups, as calculated.
    - The distribution of the observations in each group are normally distributed.
    - These groups are not paired.
    - It is obvious that the OJ supplement is more effective with the tooth growth than the VC supplement especially for the 0.5mg and 1mg dosage.
    -It is also obvious that an increase in dosage shows an increase in the tooth growth with both supplements.
    

    
