---
title: "Statistical Inference Project Part 2"
Date: Dec. 21st. 2014
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    theme: spacelab
    toc: yes
  pdf_document:
    fig_caption: yes
    highlight: tango
    number_sections: yes
    toc: yes
Author: Ahmed Assal
---
## Synopsis
This is the project for the statistical inference class.  This is the second part of the project which is an analysis of the ToothGrowth data. 

In the second portion of the class, I am going to analyze the ToothGrowth data in the R datasets package.

## Data Description    
The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

[,1]   len   numeric	 Tooth length   
[,2]	 supp	 factor	 Supplement type (VC or OJ).   
[,3]	 dose	 numeric	 Dose in milligrams.   
*(source: R Doumentation)*  

## Data Processing
### Loading packages and libraries   
```{r results='hide'}
library(ggplot2)
library(datasets)
library(reshape2)
library(dplyr)
```
### Loading dataset
```{r}
toothGrowth = tbl_df(ToothGrowth)
```

## Anaylsis
### Summarizing Data

```{r}
glimpse(toothGrowth)
summary(toothGrowth)
head(toothGrowth,3); tail(toothGrowth,3)
```

### Exploratory Data Analysis
#### Tooth length vs supplement type

Fisrt, I will explore the relaion between the supplement type and the lenth of tooth
    
```{r}
ggplot(data=toothGrowth, aes(x=len)) + 
  geom_histogram(binwidth=2) +
  facet_grid( . ~ supp) +
  xlab("Tooth Length") +
  ylab("Count")
```   

It seems that both supplements have similar effects on the length of tooth, however, the orange juice, OJ, has more profound effect on most of the teeth    

#### Tooth length vs dosage    
I will then explore the effect of the dosage quantity on the length of the tooth while examinning this effect for both of the supplements.    

```{r}
ggplot(data=toothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) + 
  geom_bar(stat="identity") +
  xlab("Dosage (mg)") +
  ylab("Tooth Length") +
  guides(fill=guide_legend(title="Supplement Type"))
```     

As the dosage increases so does the length of the tooth, however, again the orange juice, the OJ supplement has a stronger effect on the tooth length.

### Comparison of tooth growth by supp and dose   
    Did the student perform some relevant confidence intervals and/or tests?Were the results of the tests and/or intervals interpreted in the context of the problem correctly? 

## Conslusions and Assumptions        
    Did the student describe the assumptions needed for their conclusions?
